---
title: "Alcohol Consumption Data Analysis"
author: "Gopinath Achuthan"
date: "2024-09-22"
output: html_document
---

## Description of the data

This dataset measures the average alcohol consumption per person in different countries, specifically focusing on beer, spirits, and wine servings. It provides insights into drinking habits globally and can help address research questions related to cultural consumption patterns, public health implications, and economic factors surrounding alcohol production and sales. The data is stored in a CSV format, which is a delimited text file where each line represents a row of data, and the values are separated by commas.

## Reading the data into R

```{r load, echo=TRUE}

# Load the necessary package
suppressMessages(library(readr)) # for reading csv files into a dataframe
suppressMessages(suppressWarnings(library(dplyr))) # for data cleaning
suppressMessages(suppressWarnings(library(knitr))) # for displaying table

# Read the CSV file and assign it to a dataframe
url <- 'https://raw.githubusercontent.com/fivethirtyeight/data/refs/heads/master/alcohol-consumption/drinks.csv'
alcohol_data <- read_csv(url, 
                         col_names = TRUE,
                         show_col_types = FALSE)

# The col_names = TRUE argument in read_csv() indicates that the first row of the CSV file contains the column names
# Setting show_col_types = FALSE in read_csv() suppresses the display of column type information for a cleaner output

# Display the raw data
kable(head(alcohol_data, 5),
      caption = '<div style="text-align: center; color: turquoise;">Top 5 Raw Data</div>',
      align = 'c'
)

```
## Clean the data

```{r clean, echo=TRUE}

# Clean the data: Renaming columns and filtering the column by Keeping only countries with positive alcohol consumption
cleaned_data <- alcohol_data %>%
  rename(
    country_name = country,
    beer_servings_per_person = beer_servings,
    spirit_servings_per_person = spirit_servings,
    wine_servings_per_person = wine_servings,
    total_alcohol_litres_per_person = total_litres_of_pure_alcohol
  ) %>%
  filter(total_alcohol_litres_per_person > 0)

# Display the cleaned data
kable(head(cleaned_data, 5),
      caption = '<div style="text-align: center; color: turquoise;">Top 5 Cleaned Data</div>',
      align = 'c'
)

```

## Characteristics of the data

```{r characteristics, echo=TRUE}

# Create a characteristics table
column_info <- data.frame(
  Column_Number = 1:3,
  Column_Name = c("country_name", "beer_servings_per_person", "total_alcohol_litres_per_person"),
  Description = c("Name of the country", 
                  "Average beer serving sizes per person", 
                  "Average total pure alcohol litres per person")
)

# column alignment: "l" for left, "c" for center, "r" for right
alignment <- c("c", "l", "l")  # Center for column 1, left for columns 2 and 3

# Display the characteristics table using kable
kable(column_info, 
      caption = '<div style="text-align: center; color: turquoise;">Characteristics Table</div>',
      align = alignment)
```

## Subset and Summary

```{r summary, echo=TRUE}
# Create a subset of the dataframe with two selected columns
subset_data <- cleaned_data[, c("beer_servings_per_person", "total_alcohol_litres_per_person")]

# Function to calculate mode
get_mode <- function(x) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}

# Summarize the two columns
summary_stats <- data.frame(
  Min = c(
    min(subset_data$beer_servings_per_person, na.rm = TRUE),
    min(subset_data$total_alcohol_litres_per_person, na.rm = TRUE)
  ),
  Max = c(
    max(subset_data$beer_servings_per_person, na.rm = TRUE),
    max(subset_data$total_alcohol_litres_per_person, na.rm = TRUE)
  ),
  Mean = c(
    mean(subset_data$beer_servings_per_person, na.rm = TRUE),
    mean(subset_data$total_alcohol_litres_per_person, na.rm = TRUE)
  ),
  Mode = c(
    get_mode(subset_data$beer_servings_per_person),
    get_mode(subset_data$total_alcohol_litres_per_person)
  )
)

# Set row names for clarity
rownames(summary_stats) <- c("<b>Beer Servings</b>", "<b>Total Pure Alcohol</b>")

# assign summary stats to a new object
summary_result <- summary_stats

# Display the summary results using kable
kable(summary_result, 
      caption = '<div style="text-align: center; color: turquoise;">Summary Statistics</div>',
      align = 'c')
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
